---
title: "exp2_preprocess"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## import
```{r message=FALSE, results='hide'}
library(knitr)
library(tidyverse)

# fnames = list.files(path='../data/test/', pattern='ver2_beta_pilot_sandbox*', full.names=T) # sandbox
# fnames = list.files(path='../data/ver2_pilot/', pattern='MinPath_ver2_pilot_*', full.names=T) # pilot
fnames = list.files(path='../data/ver2/', pattern='MinPath_ver2*', full.names=T) # experiment
foutname = '../data/exp2_tidy'
```

## load and tidy data
``` {r}
data = do.call(rbind, lapply(fnames, function(f){read.csv(f, 
                                                          header=TRUE, 
                                                          sep=",", 
                                                          stringsAsFactors = FALSE)}))
N = nrow(data)

data = data %>%
  # assign subject id
  mutate(subject = 1:N) %>%
  # extract experiment relevant columns
  select(subject, starts_with('Answer.')) %>%
  # rename columns
  rename_at(vars(starts_with('Answer.')), list(~str_replace(., 'Answer.', '')))

data.survey = data %>%
  select('subject', 'totalBonus', starts_with('q')) %>%
  mutate_at('totalBonus', as.numeric) %>%
  mutate_at(c('q1', 'q3', 'q5', 'q6', 'q7', 'q8', 'q9', 'q10'), 
            ~factor(., levels=c('never','rarely','sometimes','often','always')))

data = data %>%
  select(-starts_with('q')) %>%
  filter(totalBonus != '{}') # didn't finish experiment
```

## check orientation group sizes
``` {r}
data = data %>%
  group_by(subject) %>%
  mutate(sideGroup = strsplit(side, ',')[[1]][1]) %>% # extract orientation on the first trial
  mutate(sideGroup = ifelse(sideGroup %in% c('east','west'), 'east/west',
                            'north/south')) %>%
  ungroup()

cat('east/west   N =', data %>% filter(sideGroup=='east/west') %>% nrow(),
    '\nnorth/south N =', data %>% filter(sideGroup=='north/south') %>% nrow())

cat('\nno counting N =', data.survey %>% filter(q1 %in% c('never', 'rarely')) %>% nrow(),
    '\ncounting    N =', data.survey %>% filter(q1 %in% c('sometimes', 'often', 'always')) %>% nrow())
```

## expand to trial level
``` {r}
# split 1d arrays
colnames1 = c('trialIndex', 'level', 'side',
              'agentIniLoc', 'goalLoc',
              'optimStep', 'totalStep', 'bonus',
              'startTime', 'startRT', 'endRT');
t1 = data %>% 
  separate_rows(colnames1, sep=',') %>% 
  select(subject, totalBonus, colnames1) # include three non-expandable columns

# split 2d arrays
colnames2 = c('subgoalLoc',
              'actionSeq', 'actionSeqStr', 'actionSeqEffect', 
              'actionSeqTime', 'actionSeqRT')
t2 = data %>%
  separate_rows(colnames2, sep='\\[|\\],\\[|\\]') %>%
  filter(actionSeq!='') %>% # remove the first and last row due to regex
  select(colnames2)

# combine t1 and t2
data = cbind(t1, t2) # bind by columns
remove(t1, t2)
```

## numeric cast and bonus-based exclusion
``` {r}
# convert some numeric columns from character columns
# reorder columns
numeric.cols = c('totalBonus', 'trialIndex', 'agentIniLoc', 'goalLoc',
                 'optimStep', 'totalStep', 'bonus', 
                 'startTime', 'startRT', 'endRT')

data = data %>%
  mutate_at(numeric.cols, as.numeric)

# !note: the original experiment encoded west/east sides as opposite, 
# labeling trials where the agent starts from the left side as the east trials instead.
# correcting that here
data = data %>%
  rename(sideDeprecated=side) %>%
  mutate(side=ifelse(sideDeprecated=='west','east',
              ifelse(sideDeprecated=='east','west',
                     sideDeprecated)))
```

## save tidy data
``` {r}
write.csv(data, paste(foutname,'.csv', sep=''))
write.csv(data.survey, paste(foutname, '_survey.csv', sep=''))
```

## Expand to step level
``` {r}
data = data %>%
  separate_rows('actionSeq', 'actionSeqStr', 'actionSeqEffect', 
                'actionSeqTime', 'actionSeqRT',
                sep=',') %>%
  group_by(subject, trialIndex) %>%
  mutate(step = 1:n()) %>%
  ungroup() %>%
  select(subject, trialIndex, level, step, everything()) %>%
  rename(action=actionSeq, actionStr=actionSeqStr,actionEffect=actionSeqEffect, 
         actionTime=actionSeqTime, actionRT=actionSeqRT) %>%
  mutate_at('actionEffect', as.logical) %>%
  mutate_at(c('actionTime', 'actionRT'), as.numeric)

data = data %>%
  # convert actions to canonical actions 
  # use west orientation as the reference orientation
  # !corrected for the mislabeled east/west sides
  mutate(canonicalUp = ((side=='west')&(actionStr=='up')) |
                       ((side=='north')&(actionStr=='right')) |
                       ((side=='east')&(actionStr=='down')) |
                       ((side=='south')&(actionStr=='left')),
         canonicalDown = ((side=='west')&(actionStr=='down')) |
                       ((side=='north')&(actionStr=='left')) |
                       ((side=='east')&(actionStr=='up')) |
                       ((side=='south')&(actionStr=='right')),
         canonicalLeft = ((side=='west')&(actionStr=='left')) |
                       ((side=='north')&(actionStr=='up')) |
                       ((side=='east')&(actionStr=='right')) |
                       ((side=='south')&(actionStr=='down')),
         canonicalRight = ((side=='west')&(actionStr=='right')) |
                       ((side=='north')&(actionStr=='down')) |
                       ((side=='east')&(actionStr=='left')) |
                       ((side=='south')&(actionStr=='up'))) %>%
  mutate(canonicalActionStr = ifelse(canonicalUp, 'up',
                              ifelse(canonicalDown, 'down',
                              ifelse(canonicalLeft, 'left',
                              ifelse(canonicalRight, 'right', 'n/a'))))) %>%
  select(-canonicalUp, -canonicalDown, -canonicalLeft, -canonicalRight)
```

## Fold back
``` {r}
data = data %>%
  nest(steps=c('step','action', 'actionStr', 'canonicalActionStr',
               'actionEffect', 'actionTime', 'actionRT'))
```

# ----------- add additional columns -----------

## filter trials
``` {r}
data = data %>%
  filter(level != 'prac')
```

## extract planning time
``` {r}
data = data %>%
  unnest(steps) %>%
  group_by(subject, trialIndex) %>%
  mutate(planningTime = actionRT[which(!is.na(actionRT))[1]]) %>%
  ungroup() %>%
  nest(steps=c('step', 'action', 'actionStr', 'canonicalActionStr',
               'actionEffect', 'actionTime', 'actionRT'))
```

## Plotting planning time distributions
``` {r}

c = sunset[1]

data %>%
  mutate(planningTime = planningTime/1000.0) %>%
  ggplot(data = .,
         mapping = aes(x=planningTime, group=subject)) +
  geom_density(color='black', fill=c, alpha=0.6) +
  scale_x_log10() +
  labs(x='Planning Time (s)')
```

## exclude ineffective steps and add some useful action columns
``` {r}
data = data %>%
  unnest(steps) %>%
  filter(actionEffect) %>%
  group_by(subject, trialIndex) %>%
  mutate(step = 1:n()) %>% # recode step index after removing ineffective steps
  mutate(upDownAction = canonicalActionStr %in% c('up','down'),
         firstRightStep = which(canonicalActionStr=='right')[1],
         lastRightStep = tail(which(canonicalActionStr=='right'),1)) %>%
  mutate(firstAction = canonicalActionStr[1],
         lastAction = ifelse(level=='base',
                             tail(canonicalActionStr, 1),
                             # the last action before subgoal is the last up/down action before the last series of horizontal actions
                             # note the index for lastRightStep: lastRightStep is a trial-level var and the data frame now is step-level
                             canonicalActionStr[tail(which(upDownAction[1:lastRightStep[1]]),1)])) %>%
  dplyr::select(-upDownAction) %>% # only a convenience column
  ungroup() %>%
  nest(steps=c('step', 'action', 'actionStr', 'canonicalActionStr',
               'actionEffect', 'actionTime', 'actionRT'))
```

## detect path start/end
``` {r}
data = data %>%
  unnest(steps) %>%
  # finally! create variables indicating the path the subject took to 
  # move the sprite initially and to approach the goal
  group_by(subject, trialIndex) %>%
  # when detecting path start, we use the last action before turning right
  # for path end, we use the last step to reach the goal/subgoal
  mutate(pathStart = ifelse(canonicalActionStr[firstRightStep-1]=='up',
                            'upper',
                     ifelse(canonicalActionStr[firstRightStep-1]=='down',
                            'lower',
                            'other')),
         pathEnd = ifelse(lastAction == 'down',
                          'upper',
                   ifelse(lastAction == 'up',
                          'lower',
                          'other')),
         pathTaken = ifelse( (pathStart=='upper')&(pathEnd=='upper'),
                             'upper',
                     ifelse( (pathStart=='lower')&(pathEnd=='lower'),
                             'lower',
                             'other'))) %>%
  # add a column indicating change of plan at the start of the trial
  # we want to compare the first action and pathStart (ver2 has no overshooting)
  mutate(pathStartChange = ifelse( ((firstAction=='down') & (pathStart=='upper')) |
                                   ((firstAction=='up') & (pathStart=='lower')),
                                   T,
                                   F)) %>%
  ungroup() %>%
  nest(steps=c('step', 'action', 'actionStr', 'canonicalActionStr',
               'actionEffect', 'actionTime', 'actionRT'))
```

## split data
### split main and ms data
``` {r}
data.main = data %>%
  filter(level %in% c('base', 'subgoal'))

data.ms = data %>%
  filter(level %in% c('multisubgoal', 'mscontrol'))
```

## base and subgoal trials

### add advantage types
``` {r}
data.main = data.main %>%
  mutate(subgoalLoc=as.numeric(subgoalLoc)) %>%
  mutate(performance = ifelse(bonus==0.03,
                              'optimal', 
                              'nonoptimal'),
         # compute myopic advantage
         myopicQuantAdv = agentIniLoc-5, # wall center is always 5
         myopicAdvPath = ifelse(myopicQuantAdv<0,  
                                'upper',
                         ifelse(myopicQuantAdv==0, 
                                'neutral', 
                                'lower')),
         # compute future advantage
         futureQuantAdv = ifelse(level=='base',
                                 goalLoc-5,
                                 subgoalLoc-5),
         futureAdvPath = ifelse(futureQuantAdv<0,
                                'upper',
                         ifelse(futureQuantAdv==0,
                                'neutral',
                                'lower')),
         # compute total advantage steps and path
         totalAdv = (myopicQuantAdv + futureQuantAdv) * 2,
         totalAdvSteps = abs(totalAdv),
         totalAdvPath = ifelse(totalAdv<0, 
                               'upper',
                        ifelse(totalAdv==0, 
                               'neutral',
                               'lower')),
         # compute trial type based on myopic/goal local advantage congruence
         trialAdvType = ifelse((myopicAdvPath=='neutral')&(futureAdvPath=='neutral'),
                               'NT',
                        ifelse((myopicAdvPath!='neutral')&(futureAdvPath=='neutral'),
                               'MA', 
                        ifelse((myopicAdvPath=='neutral')&(futureAdvPath!='neutral'),
                               'FA',
                        ifelse(((myopicAdvPath=='upper')&(futureAdvPath=='upper')) |
                               ((myopicAdvPath=='lower')&(futureAdvPath=='lower')),
                               'CA', 
                               'IA')))))

# !note: reversing advantage coding so positive=upper (more intuitive)
data.main = data.main %>%
  mutate(myopicQuantAdv=-myopicQuantAdv,
         futureQuantAdv=-futureQuantAdv,
         totalAdv=-totalAdv)
```

## save augmented data
``` {r}
saveRDS(data.main, paste(foutname,'_main.RData',sep=''))
saveRDS(data.ms, paste(foutname,'_ms.RData',sep=''))
```
