---
title: "exp2_analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# import
```{r message=FALSE, results='hide'}
library(knitr)
library(tidyverse)
library(lme4)
library(lmerTest)
library(emmeans)
library(ggeffects)
library(LaplacesDemon)
library(see)
library(patchwork)
library(rtdists)
library(parameters)
library(scales)

require(parallel)
MIN_ZRT = 0.5

select <- dplyr::select

theme_set(theme_classic() + 
          theme(text = element_text(size = 14)) +
          theme(axis.text=element_text(size=10)) +
          theme(axis.title=element_text(size=12)) +
          theme(
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            panel.background = element_rect(fill = "transparent", colour = NA),
            plot.background = element_rect(fill = "transparent", colour = NA)
          ))

finname = '../data/exp2_tidy'
plotout = '../plots/exp2/'

redblue = c('#F6B190', '#DF7C6A', '#BE364A', '#91C5DF', '#1B64AE')
```

# ----------- load augmented data -----------

``` {r}
data.main = readRDS(paste(finname,'_main.RData',sep=''))
data.survey = read.csv(paste(finname, '_survey.csv', sep=''))
```

## exclude trials
``` {r}
data.main = data.main %>%
  # trials with exceedingly long planning time (n=19)
  filter(planningTime<60000) %>%
  # remove trials solved with 4+ steps than the longer canonical solution (n=10)
  filter(totalStep <= (optimStep + totalAdvSteps + 4)) %>%
  # trials with ill-identified path start (n=180)
  filter(pathTaken!='other', pathStartChange==F)
```

```{r}
data.main = data.main %>%
  mutate(relativeAdv = abs(myopicQuantAdv) - abs(futureQuantAdv)) %>%
  # rename some trial types
  mutate(trialAdvType = ifelse(trialAdvType=='MA', 'SA-m', trialAdvType)) %>%
  mutate(trialAdvType = ifelse(trialAdvType=='FA', 'SA-f', trialAdvType)) %>%
  mutate(trialAdvType = ifelse( (trialAdvType=='IA') & (relativeAdv<0),
                               'IA-f',
                               trialAdvType)) %>%
  mutate(trialAdvType = ifelse( (trialAdvType=='IA') & (relativeAdv>0),
                               'IA-m',
                               trialAdvType)) %>%
  mutate(trialAdvType = factor(trialAdvType, level=c('NT', 'SA-m', 'SA-f', 'CA', 'IA', 'IA-m', 'IA-f')))
```

# accuracy groups
```{r}
# seprarate fit for participants with top half accuracy and bottom half accuracy
# top half optimal rate range: 0.9166667 1.0000000
# bottom half optimal rate range: 0.5384615 0.9158879
top_half = data.main %>% 
  filter(totalAdv!=0) %>%
  group_by(subject) %>%
  summarise(optimal=mean(pathStart==totalAdvPath)) %>% 
  filter(optimal >= median(optimal)) %>% 
  pull(subject)
```

## optimality and planning time in all trials by trial total advantage (path len diff)
```{r}
p1 = data.main %>%
  # filter(! subject %in% top_half) %>%
  filter(totalAdv==0) %>% # NT and IA-equal trials
  mutate(totalAdv=factor(abs(totalAdv)),
         myopicQuantAdv=factor(abs(myopicQuantAdv), levels=c(0,1,2))) %>%
  mutate(trialAdvType=as.character(trialAdvType)) %>%
  mutate(trialAdvType=ifelse(trialAdvType=='IA', paste(trialAdvType,myopicQuantAdv,sep='-'),
                             trialAdvType)) %>%
  mutate(trialAdvType=factor(trialAdvType, levels=c('NT','IA-1','IA-2'))) %>%
  mutate(pathFavorsMyopicAdv = ifelse(trialAdvType=='NT',
                                      0.5, # ipso facto
                                      as.numeric(pathStart==myopicAdvPath))) %>%
  group_by(subject, totalAdv, trialAdvType, level) %>%
  summarize(pathFavorsMyopicAdv=mean(pathFavorsMyopicAdv)) %>%
  mutate(pathFavorsMyopicAdvProbit = qnorm(pathFavorsMyopicAdv)) %>%
  mutate(pathFavorsMyopicAdvProbit = ifelse(pathFavorsMyopicAdvProbit > 3, 3,  # cap infs
                                            pathFavorsMyopicAdvProbit)) %>%
  mutate(pathFavorsMyopicAdvProbit = ifelse(pathFavorsMyopicAdvProbit < -3, -3,  # cap infs
                                            pathFavorsMyopicAdvProbit)) %>%
  ggplot(., mapping=aes(x=totalAdv, y=pathFavorsMyopicAdvProbit, 
                        color=trialAdvType, shape=trialAdvType)) +
  facet_grid(.~level) +
  stat_summary(fun.data = "mean_cl_boot",
               geom = "pointrange",
               fill = 'white',
               position = position_dodge(width=0.2),
               size = .5) +
  labs(x='Total Path Difference',
       y='Path Favors Myopic Advantage\n(probit space)') +
  scale_color_manual(labels=c('NT','IA(|mAdv|=1)','IA(|mAdv|=2)'),
                    values=c('black', redblue[4], redblue[5]),
                    name='Adv. Type') +
  scale_shape_manual(labels=c('NT','IA(|mAdv|=1)','IA(|mAdv|=2)'),
                     values=c(24,22,22),
                     name='Adv. Type') +
  coord_cartesian(ylim=c(0, 3)) +
  theme(legend.position='none')
```

```{r}
# optimal direction descriptive stats
x = data.main %>%
  filter(totalAdv!=0) %>%
  group_by(subject, level) %>%
  summarise(optimal=mean(pathStart==totalAdvPath))
t.test(x %>% filter(level=='base') %>% pull(optimal),
       x %>% filter(level=='subgoal') %>% pull(optimal),
       paired=T)
# t = 6.6057, df = 99, p-value = 1.998e-09
x %>%
  group_by(level) %>%
  summarize(xmean=mean(optimal),
            xmin=min(optimal),
            xmax=max(optimal))
#   level   xmean  xmin  xmax
# * <chr>   <dbl> <dbl> <dbl>
# 1 base    0.923 0.528     1
# 2 subgoal 0.880 0.514     1

p2 = data.main %>%
  # filter(! subject %in% top_half) %>%
  filter(totalAdv!=0) %>%
  mutate(absTotalAdv=factor(abs(totalAdv))) %>%
  group_by(subject, absTotalAdv, trialAdvType, level) %>%
  summarise(optimal=mean(pathStart==totalAdvPath)) %>%
  mutate(optimalProbit = qnorm(optimal)) %>%
  mutate(optimalProbit = ifelse(optimalProbit > 3, 3,  # cap infs
                                optimalProbit)) %>%
  mutate(optimalProbit = ifelse(optimalProbit < -3, -3,  # cap infs
                                optimalProbit)) %>%
  ggplot(., mapping=aes(x=absTotalAdv, y=optimalProbit, 
                        fill=trialAdvType, shape=trialAdvType)) +
  facet_grid(.~level) +
  stat_summary(fun.data = "mean_cl_boot",
               geom = "pointrange",
               color = 'black',
               position = position_dodge(width=0.3),
               size = .5) +
  labs(x='Total Path Difference',
      y='Optimal Direction\n(probit space)') +
  scale_fill_manual(values=c(redblue[5], redblue[2], 'white', redblue[5], redblue[2]),
                    name='Adv. Type') +
  scale_shape_manual(values=c(21,21,23,22,22),
                    name='Adv. Type') +
  coord_cartesian(ylim=c(0, 3)) +
  theme(legend.position='none')
```

```{r}
x = data.main %>%
  mutate(pt=planningTime/1000) %>%
  group_by(subject, level) %>%
  summarise(xmed=median(pt)) 
t.test(x %>% filter(level=='subgoal') %>% pull(xmed),
       x %>% filter(level=='base') %>% pull(xmed),
       paired=T)
# t = -5.8577, df = 99, p-value = 6.149e-08
x %>%
  group_by(level) %>%
  summarize(xmean=mean(xmed),
            xmin=min(xmed),
            xmax=max(xmed))
# 1	base	1.530160	0.5805	6.157
# 2	subgoal	1.912325	0.5490	9.446

p3 = data.main %>%
  # filter(! subject %in% top_half) %>%
  mutate(totalAdv=factor(abs(totalAdv)),
         myopicQuantAdv=factor(abs(myopicQuantAdv), levels=c(0,1,2))) %>%
  mutate(trialAdvType=as.character(trialAdvType)) %>%
  mutate(trialAdvType=ifelse(trialAdvType=='IA', paste(trialAdvType,myopicQuantAdv,sep='-'),
                             trialAdvType)) %>%
  mutate(trialAdvType=factor(trialAdvType, levels=c('NT','IA-1','IA-2','SA-m','SA-f','CA','IA-m','IA-f'))) %>%
  mutate(pt=planningTime/1000) %>%
  group_by(subject) %>%
  mutate(meanpt=mean(pt),
         stdpt=sd(pt)) %>%
  mutate(zpt=(pt-meanpt)/stdpt) %>%
  ungroup() %>%
  group_by(subject, level, totalAdv, trialAdvType) %>%
  summarize(meanpt=mean(meanpt),
            stdpt=mean(stdpt),
            medianzpt=median(zpt)) %>%
  ungroup() %>%
  mutate(popmeanpt=mean(meanpt), 
         popstdpt=mean(stdpt)) %>%
  mutate(projectedmedianzpt=medianzpt*popstdpt+popmeanpt) %>%
  ggplot(., mapping=aes(x=totalAdv, y=projectedmedianzpt, 
                        fill=trialAdvType, shape=trialAdvType, color=trialAdvType)) +
  facet_grid(.~level) +
  stat_summary(fun.data = "mean_cl_boot",
               geom = "pointrange",
               position = position_dodge(width=0.3),
               size = .5) +
  labs(x='Total Path Difference',
      y='Projected Median\nz(First Move Response Time) (s)') +
  scale_fill_manual(labels=c('NT','IA(|mAdv|=1)','IA(|mAdv|=2)','SA-m','SA-f','CA','IA-m','IA-f'),
                    values=c('white','white','white', redblue[5], redblue[2], 'white', redblue[5], redblue[2]),
                    name='Adv. Type') +
  scale_color_manual(labels=c('NT','IA(|mAdv|=1)','IA(|mAdv|=2)','SA-m','SA-f','CA','IA-m','IA-f'),
                     values=c('black', redblue[4], redblue[5], 'black', 'black', 'black', 'black', 'black'),
                     name='Adv. Type') +
  scale_shape_manual(labels=c('NT','IA(|mAdv|=1)','IA(|mAdv|=2)','SA-m','SA-f','CA','IA-m','IA-f'),
                     values=c(24,22,22,21,21,23,22,22),
                     name='Adv. Type') +
  coord_cartesian(ylim=c(1, 2.8)) +
  theme(legend.title=element_text(size=9), 
        legend.text=element_text(size=9),
        legend.key.size=unit(0.75, 'line'))
```

```{r, fig.width=5, fig.height=1.5}
p1 + p2 + p3 + 
  plot_annotation(tag_levels='A') +
  plot_layout(ncol=3, widths = c(1,1.5,2)) &
  theme(axis.title = element_text(size=11))
 
# ggsave(paste(plotout,'exp2.pdf', sep=''), bg='transparent')
# ggsave(paste(plotout,'exp2_top.pdf', sep=''), bg='transparent')
# ggsave(paste(plotout,'exp2_bottom.pdf', sep=''), bg='transparent')
```

## create unique labels for each of 12 unique advantage mappings
```{r}
data.fulltrialtypes = data.main %>%
  mutate(trialAdvType = ifelse(trialAdvType=='IA', 
                               paste(trialAdvType,abs(myopicQuantAdv),sep='-'),
                               paste(trialAdvType,abs(totalAdv),sep='-'))) %>%
  mutate(trialAdvType=factor(trialAdvType, levels=c('NT-0','IA-1','IA-2',
                                                    'SA-m-2','SA-f-2','IA-m-2','IA-f-2',
                                                    'CA-4','SA-m-4','SA-f-4', 'IA-m-4','IA-f-4')))
```

## optimal choice test
```{r}
fit.optim = data.fulltrialtypes %>%
  filter(totalAdv != 0) %>%
  mutate(optimal=pathStart==totalAdvPath) %>%
  glmer(optimal ~ 1 + trialAdvType * level + (1| subject),
        data=.,
        family=binomial(link='probit'),
        control=glmerControl(optimizer="bobyqa"))

fit.no.interaction = data.fulltrialtypes %>%
  filter(totalAdv != 0) %>%
  mutate(optimal=pathStart==totalAdvPath) %>%
  glmer(optimal ~ 1 + trialAdvType + level + (1| subject),
        data=.,
        family=binomial(link='probit'))

fit.adv = data.fulltrialtypes %>%
  filter(totalAdv != 0) %>%
  mutate(optimal=pathStart==totalAdvPath) %>%
  glmer(optimal ~ 1 + trialAdvType + (1| subject),
        data=.,
        family=binomial(link='probit'))

fit.level = data.fulltrialtypes %>%
  filter(totalAdv != 0) %>%
  mutate(optimal=pathStart==totalAdvPath) %>%
  glmer(optimal ~ 1 + level + (1| subject),
        data=.,
        family=binomial(link='probit'))

anova(fit.no.interaction, fit.optim)
# Data: .
# Models:
# fit.no.interaction: optimal ~ 1 + trialAdvType + level + (1 | subject)
# fit.optim: optimal ~ 1 + trialAdvType * level + (1 | subject)
#                    Df    AIC    BIC  logLik deviance  Chisq Chi Df Pr(>Chisq)
# fit.no.interaction 11 6108.1 6188.1 -3043.0   6086.1                         
# fit.optim          19 6115.9 6254.1 -3038.9   6077.9 8.1873      8     0.4154
anova(fit.no.interaction, fit.level)
# fit.null: optimal ~ 1 + level + (1 | subject)
# fit.no.interaction: optimal ~ 1 + trialAdvType + level + (1 | subject)
#                    Df    AIC    BIC  logLik deviance  Chisq Chi Df Pr(>Chisq)    
# fit.null            3 6633.9 6655.7 -3313.9   6627.9                             
# fit.no.interaction 11 6108.1 6188.1 -3043.0   6086.1 541.79      8  < 2.2e-16 ***
anova(fit.no.interaction, fit.adv)
# fit.adv: optimal ~ 1 + trialAdvType + (1 | subject)
# fit.no.interaction: optimal ~ 1 + trialAdvType + level + (1 | subject)
#                    Df    AIC    BIC logLik deviance  Chisq Chi Df Pr(>Chisq)    
# fit.adv            10 6168.0 6240.7  -3074   6148.0                             
# fit.no.interaction 11 6108.1 6188.1  -3043   6086.1 61.927      1  3.563e-15 ***

# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

# [1] "SA-m-2" "SA-f-2" "IA-m-2" "IA-f-2" "CA-4"   "SA-m-4" "SA-f-4" "IA-m-4" "IA-f-4"

emms = emmeans(fit.no.interaction, ~trialAdvType)

contrasts = list(sa_vs_ca=c(0,0,0,0, -1,.5,.5,0,0),
                 ia_vs_ca=c(0,0,0,0, -1,0,0,.5,.5),
                 sa_vs_ia=c(-.25,-.25,.25,.25, 0,-.25,-.25,.25,.25),
                 m_vs_f=c(-.25,.25,-.25,.25, 0,-.25,.25,-.25,.25),
                 ta2_vs_ta4=c(-1/4,-1/4,-1/4,-1/4, 1/5,1/5,1/5,1/5,1/5))

contrast(emms, contrasts, adjust='bonferroni')
#  contrast   estimate     SE  df z.ratio p.value
#  sa_vs_ca     -0.105 0.0903 Inf  -1.165 1.0000 
#  ia_vs_ca     -0.564 0.0846 Inf  -6.662 <.0001 
#  sa_vs_ia     -0.501 0.0404 Inf -12.404 <.0001 
#  m_vs_f       -0.326 0.0403 Inf  -8.104 <.0001 
#  ta2_vs_ta4    0.588 0.0387 Inf  15.221 <.0001 
# 
# Results are averaged over the levels of: level 
# P value adjustment: bonferroni method for 5 tests 

# get grouped emmeans for each contrast
group=contrasts['ia_vs_ca'][[1]]

refg = ref_grid(fit.no.interaction) %>%
  add_grouping('group', 'trialAdvType', group)

emmeans(refg, ~group)
```

## rt test
```{r}
fit.rt = data.fulltrialtypes %>%
  mutate(pt=planningTime/1000) %>%
  group_by(subject) %>%
  mutate(meanpt=mean(pt),
         stdpt=sd(pt)) %>%
  mutate(zpt=(pt-meanpt)/stdpt) %>%
  ungroup() %>%
  group_by(subject, level, trialAdvType) %>%
  summarize(meanpt=mean(meanpt),
            stdpt=mean(stdpt),
            medianzpt=median(zpt)) %>%
  ungroup() %>%
  lm(medianzpt ~ 1 + trialAdvType * level,
     data = .)

fit.rt %>% summary() # F-statistic: 27.57 on 23 and 2376 DF,  p-value: < 2.2e-16
anova(fit.rt)
# Analysis of Variance Table
# 
# Response: medianzpt
#                      Df  Sum Sq Mean Sq F value Pr(>F)    
# trialAdvType         11  59.158  5.3780 48.1667 <2e-16 ***
# level                 1  10.125 10.1253 90.6838 <2e-16 ***
# trialAdvType:level   11   1.515  0.1378  1.2338 0.2584    
# Residuals          2376 265.291  0.1117                   
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

levels(data.fulltrialtypes$trialAdvType) 
#  [1] "NT-0"   "IA-1"   "IA-2"   "SA-m-2" "SA-f-2" "IA-m-2" "IA-f-2" "CA-4"   "SA-m-4" "SA-f-4" "IA-m-4" "IA-f-4"

emms = emmeans(fit.rt, ~trialAdvType | level)
contrasts = list(nt_vs_ia=c(-1,0.5,0.5, 0,0,0,0, 0,0,0,0,0),
                 sa_vs_ca=c(0,0,0, 0,0,0,0, -1,.5,.5,0,0),
                 ia_vs_ca=c(0,0,0, 0,0,0,0, -1,0,0,.5,.5),
                 sa_vs_ia=c(0,0,0, -.25,-.25,.25,.25, 0,-.25,-.25,.25,.25),
                 m_vs_f=c(0,0,0, -.25,.25,-.25,.25, 0,-.25,.25,-.25,.25),
                 ta0_vs_ta2=c(-1/3,-1/3,-1/3, 1/4,1/4,1/4,1/4, 0,0,0,0,0),
                 ta0_vs_ta4=c(-1/3,-1/3,-1/3, 0,0,0,0, 1/5,1/5,1/5,1/5,1/5),
                 ta2_vs_ta4=c(0,0,0, -1/4,-1/4,-1/4,-1/4, 1/5,1/5,1/5,1/5,1/5))

contrast(emms, contrasts, adjust='bonferroni')
# level = base:
#  contrast   estimate     SE   df t.ratio p.value
#  nt_vs_ia     0.1405 0.0409 2376   3.432 0.0049 
#  sa_vs_ca     0.0675 0.0409 2376   1.650 0.7931 
#  ia_vs_ca     0.2389 0.0409 2376   5.836 <.0001 
#  sa_vs_ia     0.1665 0.0236 2376   7.046 <.0001 
#  m_vs_f       0.1426 0.0236 2376   6.035 <.0001 
#  ta0_vs_ta2  -0.1668 0.0255 2376  -6.538 <.0001 
#  ta0_vs_ta4  -0.3161 0.0244 2376 -12.953 <.0001 
#  ta2_vs_ta4  -0.1492 0.0224 2376  -6.658 <.0001 
# 
# level = subgoal:
#  contrast   estimate     SE   df t.ratio p.value
#  nt_vs_ia    -0.0353 0.0409 2376  -0.863 1.0000 
#  sa_vs_ca     0.0630 0.0409 2376   1.539 0.9907 
#  ia_vs_ca     0.2201 0.0409 2376   5.378 <.0001 
#  sa_vs_ia     0.1302 0.0236 2376   5.509 <.0001 
#  m_vs_f       0.0953 0.0236 2376   4.035 0.0005 
#  ta0_vs_ta2  -0.1815 0.0255 2376  -7.111 <.0001 
#  ta0_vs_ta4  -0.3306 0.0244 2376 -13.546 <.0001 
#  ta2_vs_ta4  -0.1491 0.0224 2376  -6.651 <.0001 

# get grouped emmeans for each contrast
group=contrasts['m_vs_f'][[1]]

refg = ref_grid(fit.rt) %>%
  add_grouping('group', 'trialAdvType', group)

emmeans(refg, ~group)
```

## competing myopic and future advantages
```{r}
data.ia = data.main %>%
  filter(trialAdvType %in% c('IA','IA-m','IA-f'))

p.ia.main = data.ia %>%
  mutate(myopicGoalSameSide = ifelse(level=='base', '',
                                    ((myopicAdvPath=='upper')&(goalLoc==1)) |
                                    ((myopicAdvPath=='lower')&(goalLoc==9)))) %>%
  mutate(trialType = ifelse(level=='base', 'base',
                     ifelse(myopicGoalSameSide, 'subgoal (goal on myopic side)', 'subgoal (goal on future side)'))) %>%
  group_by(subject, relativeAdv, trialType) %>%
  summarize(pathFavorsMyopicAdv = mean(pathStart==myopicAdvPath)) %>%
  mutate(pathFavorsMyopicAdvProbit = qnorm(pathFavorsMyopicAdv)) %>%
  mutate(pathFavorsMyopicAdvProbit = ifelse(pathFavorsMyopicAdvProbit > 3, 3,  # cap infs
                                            pathFavorsMyopicAdvProbit)) %>%
  mutate(pathFavorsMyopicAdvProbit = ifelse(pathFavorsMyopicAdvProbit < -3, -3,  # cap infs
                                            pathFavorsMyopicAdvProbit)) %>%
  ggplot(data=., mapping=aes(x=relativeAdv, y=pathFavorsMyopicAdvProbit,
                             fill=trialType, shape=trialType)) +
  geom_hline(aes(yintercept=0), alpha=0.5, linetype='dashed') +
  stat_summary(fun.data = "mean_cl_boot",
               geom = "pointrange",
               color = "black",
               size = .5,
               position = position_jitter(width=0.15)) +
  scale_color_manual(values=c(redblue[4],redblue[5],redblue[5]), name='Trial Type') +
  scale_fill_manual(values=c(redblue[4],redblue[5],redblue[5]), name='Trial Type') +
  scale_shape_manual(values=c(21,22,23), name='Trial Type') +
  labs(x = 'Myopic/Future Relative Advantage',
       y = 'Path Favors Myopic Advantage\n(probit space)') +
  coord_cartesian(ylim=c(-2.7,2.7),
                  xlim=c(-2.3,2.3)) +
  theme(legend.position=c(0.75, 0.25))
```

## Drift-diffusion model
```{r}
data.fit = data.main %>%
  group_by(subject) %>%
  mutate(rt=scale(planningTime/1000)) %>%
  # compensate smallest rt for rt's smaller than min_rt since ddiffusion wants positive t0
  mutate(rt=rt-min(rt)+MIN_ZRT) %>%
  ungroup() %>%
  filter(grepl('IA', trialAdvType)) %>%
  mutate(pathFavorsMyopicAdv = as.numeric(pathStart==myopicAdvPath)) %>%
  mutate(response = ifelse(pathFavorsMyopicAdv==T, 'upper', 'lower')) %>%
  # use absolute values for fitting purpose, essentially folding over up and down cases
  mutate(myopicQuantAdv=abs(myopicQuantAdv),
         futureQuantAdv=abs(futureQuantAdv)) %>%
  mutate(myopicGoalSameSide = ((myopicAdvPath=='upper')&(goalLoc==1)) |
                              ((myopicAdvPath=='lower')&(goalLoc==9))) %>%
  mutate(goalSide = ifelse(level=='base', 'na', 
                           ifelse(myopicGoalSameSide, 'myopic', 'future'))) %>%
  select(subject, level, goalSide, myopicQuantAdv, futureQuantAdv, rt, response)

myopicQuantAdv = data.fit %>% 
  select(myopicQuantAdv, futureQuantAdv) %>% 
  unique() %>% pull(myopicQuantAdv)
futureQuantAdv = data.fit %>% 
  select(myopicQuantAdv, futureQuantAdv) %>% 
  unique() %>% pull(futureQuantAdv)
all_subjects = data.fit %>% pull(subject) %>% unique()
```

```{r}
# ddiffusion uses 'v' to denote drift rate, i'm using d...

compute_drift_rate = function(pars, ma, fa, trial, goalSide) {
    # extract drift rate from pars here otherwise the clause gets too clumsy
    if(trial=='base') {
        return(ma*pars['md']-fa*pars['fd'])
    } else if (trial=='subgoal') {
        if (goalSide=='myopic') {
          gd = pars['gd']
        } else if (goalSide=='future') {
          gd = -pars['gd']
        }
        if ('dsg' %in% names(pars)) {
            return(ma*(pars['md']+pars['dsg'])-fa*(pars['fd']+pars['dsg'])+gd)
        } else if ('mdsg' %in% names(pars)) {
            return(ma*(pars['md']+pars['mdsg'])-fa*(pars['fd']+pars['fdsg'])+gd)
        } else if ('dpsg' %in% names(pars)) {
            return(ma*(pars['md']*pars['dpsg'])-fa*(pars['fd']*pars['dpsg'])+gd)
        } else {
            return(ma*pars['md']-fa*pars['fd']+gd)
        }
    }
}

diffusion_objective <- function(pars, data) {
    # loop over all 6 unique IA trials
    densities = vector('numeric', length(data$rt))
    for (i in 1:length(myopicQuantAdv)) {
        ma = myopicQuantAdv[i]
        fa = futureQuantAdv[i]
        # fit base trials
        subset_mask = (data$level=='base') & (data$myopicQuantAdv==ma) & (data$futureQuantAdv==fa)
        x = data %>% filter(subset_mask)
        densities[subset_mask] = ddiffusion(rt=x$rt, response=x$response,
                                            t0=pars['t0'],
                                            a=pars['a'],
                                            v=compute_drift_rate(pars, ma, fa, 'base', 'none'),
                                            sv=if ('sd' %in% names(pars)) pars['sd']
                                            else 0,
                                            z=if('z' %in% names(pars)) pars['z']
                                            else 0.5*pars['a'],
                                            sz=if ('sz' %in% names(pars)) pars['sz']
                                            else 0)
        # fit subgoal trials
        subset_mask = (data$level=='subgoal') & (data$goalSide=='myopic') & (data$myopicQuantAdv==ma) & (data$futureQuantAdv==fa)
        x = data %>% filter(subset_mask)
        densities[subset_mask] = ddiffusion(rt=x$rt, response=x$response,
                                            t0=if ("t0sg" %in% names(pars)) pars["t0sg"]
                                            else pars['t0'],
                                            a=pars['a'],
                                            v=compute_drift_rate(pars, ma, fa, 'subgoal', 'myopic'),
                                            sv=if ('sd' %in% names(pars)) pars['sd']
                                            else 0,
                                            z=if('z' %in% names(pars)) pars['z']
                                            else 0.5*pars['a'],
                                            sz=if ('sz' %in% names(pars)) pars['sz']
                                            else 0)
        subset_mask = (data$level=='subgoal') & (data$goalSide=='future') & (data$myopicQuantAdv==ma) & (data$futureQuantAdv==fa)
        x = data %>% filter(subset_mask)
        densities[subset_mask] = ddiffusion(rt=x$rt, response=x$response,
                                            t0=if ("t0sg" %in% names(pars)) pars["t0sg"]
                                            else pars['t0'],
                                            a=pars['a'],
                                            v=compute_drift_rate(pars, ma, fa, 'subgoal', 'future'),
                                            sv=if ('sd' %in% names(pars)) pars['sd']
                                            else 0,
                                            z=if('z' %in% names(pars)) pars['z']
                                            else 0.5*pars['a'],
                                            sz=if ('sz' %in% names(pars)) pars['sz']
                                            else 0)
    }
    if (any(densities==0)) return(1e6)
    return(-sum(log(densities)))
}

get_init_vals = function(model_spec, data) {
    # model_spec: a list of parameters, must include ['t0','a','v']
    init_ll=1e6
    while(init_ll>=1e6){ # find a valid set of initial value
        pars = c() # parameters
        lower = c() # lower bounds
        # t0: irreducible nondecision time
        if('t0' %in% model_spec) {pars=c(pars, t0=runif(1, 0, MIN_ZRT))
        lower=c(lower, 0)}
        # a: decision bound
        if('a' %in% model_spec) {pars=c(pars, a=runif(1, 0, 1))
        lower=c(lower, 0)}
        # d: drift rate (but argument v in ddiffusion)
        if('d' %in% model_spec) {pars=c(pars, d=runif(1, 0, 1))
        lower=c(lower, -Inf)}
        if('md' %in% model_spec) {pars=c(pars, md=runif(1, 0, 1))
        lower=c(lower, -Inf)}
        if('fd' %in% model_spec) {pars=c(pars, fd=runif(1, 0, 1))
        lower=c(lower, -Inf)}
        # sd: across-trial-variability of drift rate (sv in ddiffusion))
        if('sd' %in% model_spec) {pars=c(pars, sd=runif(1, 0, 1))
        lower=c(lower, 0)}
        # z: starting point
        if('z' %in% model_spec) {pars=c(pars, z=runif(1, 0, 1))
        lower=c(lower, 0)}
        # sz: across-trial-variability of starting point
        if('sz' %in% model_spec) {pars=c(pars, sz=runif(1, 0, 1))
        lower=c(lower, 0)}
        
        # subgoal variants
        if('t0sg' %in% model_spec) {pars=c(pars, t0sg=runif(1, 0, MIN_ZRT))
        lower=c(lower, 0)}
        if('dsg' %in% model_spec) {pars=c(pars, dsg=runif(1, 0, 1))
        lower=c(lower, -Inf)}
        if('mdsg' %in% model_spec) {pars=c(pars, mdsg=runif(1, 0, 1))
        lower=c(lower, -Inf)}
        if('fdsg' %in% model_spec) {pars=c(pars, fdsg=runif(1, 0, 1))
        lower=c(lower, -Inf)}
        if('dpsg' %in% model_spec) {pars=c(pars, dpsg=runif(1, 0, 1))
        lower=c(lower, 0)}
        if('gd' %in% model_spec) {pars=c(pars, gd=runif(1, 0, 1))
        lower=c(lower, 0)}
        
        init_ll = diffusion_objective(pars, data)
    }
    return (list(pars=pars, lower=lower))
}

fit_diffusion_model = function(model_spec, train_data, test_data){
    init_vals = get_init_vals(model_spec, train_data)
    fit = nlminb(init_vals$pars, 
                 diffusion_objective, 
                 data=train_data, 
                 lower=init_vals$lower)
    fit$test_objective = diffusion_objective(pars=fit[[1]], data=test_data)
    return(fit)
}

find_best_fit = function(model_spec, train_data, test_data, nfit) {
  fits = mclapply(1:nfit, function(x) {fit_diffusion_model(model_spec, 
                                                            train_data, 
                                                            test_data)}, 
                  mc.cores=2)
  # flag if any training objective is 1e6
  train_objectives = unlist(lapply(fits, '[[', 2))
  if (sum(train_objectives==1e6)>0) {cat('uh oh')} # this isn't good
  # find minimum train objective out of all fits
  train_objective_argmin = which(train_objectives==min(train_objectives))
  if (length(train_objective_argmin)>1) {train_objective_argmin=train_objective_argmin[1]} #ties
  best_fit = fits[[train_objective_argmin]]
  best_fit = as.data.frame(t(unlist(best_fit)))
  colnames(best_fit) <- sub("par.", "", colnames(best_fit))
  return(best_fit)
}

cross_val_paired = function(i, model_spec, data, nfit) {
    # performs one round of monte-carlo cross validation
    test_subjects = cv_folds[,i]
    train_data = data %>% filter(! subject %in% test_subjects)
    test_data = data %>% filter(subject %in% test_subjects)
    best_fit = find_best_fit(model_spec,
                             train_data=train_data,
                             test_data=test_data,
                             nfit=nfit)
    best_fit$test_subjects = list(test_subjects)
    return (best_fit)
}
```

```{r}
# all possible params: c('t0','a','d','md','fd','sd','z','sz','t0sg','dsg','mdsg','fdsg','dpsg','gd')

tophalf = data.fit %>% filter(subject %in% top_half)
bottomhalf = data.fit %>% filter(! subject %in% top_half)

# generate 200 CV fold used commonly across different models
nfold = 200
ntest = 40 # 15
cv_folds = as.data.frame(replicate(nfold, sample(all_subjects, ntest)))
# cv_folds = as.data.frame(replicate(nfold, sample(unique(tophalf$subject), ntest)))
# cv_folds = as.data.frame(replicate(nfold, sample(unique(bottomhalf$subject), ntest)))
# saveRDS(cv_folds, 'exp2_cv_fold.RData')
# cv_folds = readRDS('exp2_cv_fold_bottomhalf.RData')

############
# paired   #
############
model_specs = list(
                   # c('t0','a','md','fd','sz','sd','gd'),
                   # c('t0','a','md','fd','sz','sd','dsg','gd'),
                   # c('t0','a','md','fd','sz','sd','mdsg','fdsg','gd'),
                   # c('t0','a','md','fd','sz','sd','dpsg','gd'),
                   # c('t0','a','md','fd','sz','sd','t0sg','gd'),
                   # c('t0','a','md','fd','sz','sd','t0sg','dsg','gd'),
                   c('t0','a','md','fd','sz','sd','t0sg','mdsg','fdsg','gd')
                   # c('t0','a','md','fd','sz','sd','t0sg','dpsg','gd','gd')
                  )

for (i in 1:length(model_specs)) {
    spec = model_specs[[i]]
    model_out = paste('./exp2_diffusion_revision/fixed_cv/exp2_fits_',
                      paste(spec, collapse='_'),
                      '.RData', sep='')
    if(! file.exists(model_out)) {
        cat(paste('fitting and saving to ',model_out,'......\n',sep=''))
        t0 = Sys.time()
        fits = mclapply(1:nfold, function(i) {cross_val_paired(i, spec, data.fit, nfit=10)}, 
                        mc.cores=2)
        t1 = Sys.time()
        cat(t1-t0)
        saveRDS(fits, model_out)
    }
}
```

### analyze fits
```{r, warning=F}
read_fit = function(fname) {
  fit = readRDS(fname)
  fit = do.call(rbind.data.frame, fit) %>% mutate_all(as.character)
  fit = fit %>% 
      dplyr::select(-test_subjects) %>%
      mutate(fold=1:n(),
             model=str_sub(tail(str_split(fname,'/')[[1]],1),
                           start=11,end=-7))
  return(fit)
}

paired_fits = list.files(path='./exp2_diffusion_revision/fixed_cv', 
                      pattern='exp2_fits_t0_*', full.names=T)
paired_fits = do.call(bind_rows, lapply(paired_fits, read_fit)) %>%
  mutate_at(c('t0','a','md','fd','sz','sd','t0sg','dsg','mdsg','fdsg','dpsg','gd',
              'objective','test_objective'), as.numeric) %>%
  mutate(model=str_replace(model, 't0_a_md_fd_sz_sd', 'basemodel')) %>%
  mutate(model=factor(model, levels=c('basemodel',
                                      'basemodel_gd',
                                      'basemodel_dsg_gd',
                                      'basemodel_mdsg_fdsg_gd',
                                      'basemodel_dpsg_gd',
                                      'basemodel_t0sg_gd',
                                      'basemodel_t0sg_dsg_gd',
                                      'basemodel_t0sg_mdsg_fdsg_gd',
                                      'basemodel_t0sg_dpsg_gd')))

paired_fits = list.files(path='./exp2_diffusion_revision/fixed_cv', 
                      pattern='exp2_fits_tophalf*', full.names=T)
paired_fits = do.call(bind_rows, lapply(paired_fits, read_fit)) %>%
  mutate_at(c('t0','a','md','fd','sz','sd','dpsg','gd',
              'objective','test_objective'), as.numeric) %>%
  mutate(model=str_replace(model, 'tophalf_', '')) %>%
  mutate(model=str_replace(model, 't0_a_md_fd_sz_sd', 'basemodel'))

paired_fits = list.files(path='./exp2_diffusion_revision/fixed_cv', 
                      pattern='exp2_fits_bottomhalf*', full.names=T)
paired_fits = do.call(bind_rows, lapply(paired_fits, read_fit)) %>%
  mutate_at(c('t0','a','md','fd','sz','sd','dpsg','gd',
              'objective','test_objective'), as.numeric) %>%
  mutate(model=str_replace(model, 'bottomhalf_', '')) %>%
  mutate(model=str_replace(model, 't0_a_md_fd_sz_sd', 'basemodel'))
```

```{r}
data.plot = paired_fits %>%
  filter(model %in% c('basemodel','basemodel_gd','basemodel_dpsg_gd','basemodel_t0sg_gd','basemodel_t0sg_dpsg_gd')) %>%
  mutate(t0=ifelse(grepl('t0sg',model),'different','same')) %>%
  mutate(t0=factor(t0, levels=c('same','different'))) %>%
  mutate(modelname=ifelse(grepl('dpsg',model), 'shared proportional','none')) %>%
  mutate(modelname=factor(modelname, levels=c('none','shared proportional'))) %>%
  # reference all test objectives against the base model (t0_a_md_fd_sz_sd)
  # to reduce variance over CV fold
  group_by(model) %>%
  mutate(test_objective=test_objective-paired_fits %>% filter(model=='basemodel') %>% pull(test_objective)) %>%
  ungroup()

p.model.comp.full = ggplot() +
  stat_summary(data=data.plot, 
               mapping=aes(x=t0, y=test_objective, color=modelname, fill=modelname),
               fun='mean',
               geom='bar',
               width=0.85,
               position=position_dodge(width=.9)) +
  stat_summary(data=data.plot, 
               mapping=aes(x=t0, y=test_objective, fill=modelname),
               fun.data='mean_cl_boot',
               geom='errorbar',
               position=position_dodge(width=.9)) +
  stat_summary(data=data.plot %>% filter(model=='basemodel'), 
               mapping=aes(x=0.2, y=test_objective, fill=modelname),
               fun.data='mean_cl_boot',
               geom='errorbar',
               width=0.4,
               position=position_dodge(width=.9)) +
  annotate('text', label='*', x=1.23, y=0.4, size=6, hjust=0.5) +
  coord_cartesian(ylim=c(-40,2), xlim=c(0.5,2)) +
  geom_text(mapping=aes(x=0.2, y=0, label='baseline sNLL = 4013.10 (SD=114.45) '),
            hjust=1, va='right', angle=90, size=2.7, color='black') +
  geom_text(mapping=aes(x=0.2, y=1.6, label='without gd'),
            size=3.6, color='black') +
  geom_text(mapping=aes(x=0.5, y=1.6, label='|  with gd'),
            hjust=0, size=3.6, color='black') +
  scale_color_manual(values=c(redblue[3], redblue[3])) +
  scale_fill_manual(values=c('white', redblue[3])) +
  labs(y='Change in Test Data sNLL\nfrom Baseline', x='                    t0 in base trials and subgoal trials',
       fill=expression(paste('Changes to ',italic('md'),' and ',italic('fd'),' in Subgoal Trials')), 
       color=expression(paste('Changes to ',italic('md'),' and ',italic('fd'),' in Subgoal Trials'))) +
  guides(fill=guide_legend(ncol=4), color=guide_legend(ncol=4)) +
  theme(legend.position=c(0.5, 0.1),
        axis.text.x = element_text(vjust=0))
```

```{r}
levels(paired_fits$model)
# [1] "basemodel"                   "basemodel_gd"                "basemodel_dsg_gd"            "basemodel_mdsg_fdsg_gd"     
# [5] "basemodel_dpsg_gd"           "basemodel_t0sg_gd"           "basemodel_t0sg_dsg_gd"       "basemodel_t0sg_mdsg_fdsg_gd"
# [9] "basemodel_t0sg_dpsg_gd" 
model_comp = lmer(test_objective~1+model+(1|fold), paired_fits)
emms=emmeans(model_comp, ~model)

contrasts = list(none_vs_gd=c(1,-1,0,0,0,0,0,0,0),
                 same_vs_dsg=c(0,1,-1,0,0,1,-1,0,0),
                 same_vs_mdfdsg=c(0,1,0,-1,0,1,0,-1,0),
                 same_vs_dpsg=c(0,1,0,0,-1,1,0,0,-1),
                 dsg_vs_mdfdsg=c(0,0,1,-1,0,0,1,-1,0),
                 dsg_vs_dpsg=c(0,0,1,0,-1,0,1,0,-1),
                 mdfdsg_vs_dpsg=c(0,0,0,1,-1,0,0,1,-1)
                 )
contrast(emms, contrasts, adjust='bonferroni')
#  contrast       estimate    SE   df t.ratio p.value
#  none_vs_gd       12.615 0.470 1592  26.854  <.0001
#  same_vs_dsg      26.272 0.664 1592  39.545  <.0001
#  same_vs_mdfdsg   32.359 0.664 1592  48.707  <.0001
#  same_vs_dpsg     33.284 0.664 1592  50.099  <.0001
#  dsg_vs_mdfdsg     6.087 0.664 1592   9.162  <.0001
#  dsg_vs_dpsg       7.012 0.664 1592  10.554  <.0001
#  mdfdsg_vs_dpsg    0.925 0.664 1592   1.392  1.0000
# 
# Degrees-of-freedom method: kenward-roger 
# P value adjustment: bonferroni method for 7 tests 

contrasts = list(not0sg_vs_t0sg=c(0,0,0,0,1,0,0,0,-1))
contrast(emms, contrasts, adjust='bonferroni')
 # contrast       estimate   SE   df t.ratio p.value
 # not0sg_vs_t0sg    0.704 0.47 1592   1.498  0.1344
```

```{r, fig.width=4.5, fig.height=3}
# group model estimated md/fd ratio
x = paired_fits %>% filter(model=='basemodel_dpsg_gd')
mean(x$md/x$fd) # 1.215447 # top: 1.166132 # bottom: 1.303483
sd(x$md/x$fd) # 0.03429224 # top: 0.02463593 # bottom: 0.05758869

params=c('t0','a','md','fd','gd','sz','sd','p')
data.plot = paired_fits %>%
  filter(model=='basemodel_dpsg_gd') %>%
  rename(p=dpsg) %>%
  pivot_longer(params, names_to='param', values_to='value') %>%
  mutate(param=factor(param, levels=params))

p1 = data.plot %>%
  filter(param %in% c('t0')) %>%
  ggplot(data=., mapping=aes(x=1, y=value)) +     
  geom_point(position=position_jitter(width=.2),
             alpha=0.05) +
  stat_summary(fun.data = "mean_cl_boot",
               geom = "pointrange",
               size=0.3) +
  facet_grid(.~param) +
  coord_cartesian(ylim=c(0.4,0.5)) +
  labs(y='Estimate')

p2 = data.plot %>%
  filter(param %in% c('a','md','fd','gd','sz','sd')) %>%
  ggplot(data=., mapping=aes(x=1, y=value)) +     
  geom_point(position=position_jitter(width=.2),
             alpha=0.05) +
  stat_summary(fun.data = "mean_cl_boot",
               geom = "pointrange",
               size=0.3) +
  facet_grid(.~param) +
  coord_cartesian(ylim=c(0,3.5)) +
  theme(axis.title.y=element_blank())

p3 = data.plot %>%
  filter(param %in% c('p')) %>%
  ggplot(data=., mapping=aes(x=1, y=value)) +     
  geom_point(position=position_jitter(width=.2),
             alpha=0.05) +
  stat_summary(fun.data = "mean_cl_boot",
               geom = "pointrange",
               size=0.3) +
  facet_grid(.~param) +
  coord_cartesian(ylim=c(0.65,0.8)) +
  theme(axis.title.y=element_blank())

p.model.params = wrap_plots(p1 + p2 + p3 +
  plot_layout(ncol=3, widths=c(0.14,1,0.14)) &
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.x=element_blank()))

x = paired_fits%>%filter(model=='basemodel_dpsg_gd')%>%pull(gd)
mean(x) # 0.3104737 # top: 0.08209403 # bottom: 0.6799137
sd(x) # 0.05722342 # top: 0.03376332 # bottom: 0.116323
```

```{r}
model_spec = c('t0','a','md','fd','sz','sd','dpsg','gd')
best_fit = paired_fits %>%
  filter(model=='basemodel_dpsg_gd') %>%
  filter(test_objective==min(test_objective)) %>%
  select(model_spec)
# > best_fit
#          t0        a       md      fd       sz       sd      dpsg        gd
# 1 0.4716947 3.044756 1.484185 1.23733 2.341309 1.490817 0.6971697 0.3042459
# top
# > best_fit
#         t0        a       md       fd       sz       sd      dpsg         gd
# 1 0.457224 3.103193 1.611869 1.391338 2.209979 1.088728 0.6821048 0.08316488
# bottom
# > best_fit
#          t0        a       md       fd       sz       sd      dpsg        gd
# 1 0.4560321 3.029395 1.353855 1.031458 2.057995 1.930902 0.7296701 0.5004466

compute_drift_rate2 = function(pars, ma, fa, trial, goalSide) {
  # extract drift rate from pars here otherwise the clause gets too clumsy
  if(trial=='base') {
    return(ma*pars$md-fa*pars$fd)
  } else if (trial=='subgoal') {
    if (goalSide=='myopic') {
      gd = pars$gd
    } else if (goalSide=='future') {
      gd = -pars$gd
    }
    if ('dsg' %in% names(pars)) {
      return(ma*(pars$md+pars$dsg)-fa*(pars$fd+pars$dsg)+gd)
    } else if ('mdsg' %in% names(pars)) {
      return(ma*(pars$md+pars$mdsg)-fa*(pars$fd+pars$fdsg)+gd)
    } else if ('dpsg' %in% names(pars)) {
      return(ma*(pars$md*pars$dpsg)-fa*(pars$fd*pars$dpsg)+gd)
    } else {
      return(ma*pars$md-fa*pars$fd+gd)
    }
  }
}

sample_fitted_dist = function(model_spec, pars, n_sample, trial, goalSide) {
  out = map(1:length(myopicQuantAdv), 
             function(i) {
               ma=myopicQuantAdv[i]
               fa=futureQuantAdv[i]
               x=rdiffusion(n_sample, 
                            t0=if (trial=='base') pars$t0
                               else if ("t0sg" %in% names(pars)) pars$t0sg
                               else pars$t0, 
                            a=pars$a,
                            v=compute_drift_rate2(pars, ma, fa, trial, goalSide),
                            sv=if('sd' %in% names(pars)) pars$sd else 0,
                            z=if('z' %in% names(pars)) pars$z else 0.5*pars$a,
                            sz=if ('sz' %in% names(pars)) pars$sz else 0) %>%
                   mutate(ma=ma,fa=fa, level=trial)
               return(x)
             })
  out = bind_rows(out)
  return(out)
}

n_sample=50000
sampled_fit_dist = sample_fitted_dist(model_spec, best_fit, n_sample, 'base') %>%
  bind_rows(sample_fitted_dist(model_spec, best_fit, n_sample, 'subgoal', 'myopic')) %>%
  bind_rows(sample_fitted_dist(model_spec, best_fit, n_sample, 'subgoal', 'future'))
```

#### descriptive stats for empirical and fitted drift model RT responses
```{r}
data.fit.plot = data.fit %>%
  filter(! subject %in% top_half) %>%
  rename(ma=myopicQuantAdv, fa=futureQuantAdv) %>%
  mutate(mafa=paste('mAdv=',ma,', ','fAdv=',fa,sep='')) %>%
  mutate(mafa=factor(mafa, levels=c('mAdv=1, fAdv=1','mAdv=2, fAdv=2',
                                    'mAdv=2, fAdv=1','mAdv=1, fAdv=2',
                                    'mAdv=3, fAdv=1','mAdv=1, fAdv=3'))) %>%
  mutate(correct_response = ifelse(ma>fa, 'upper',
                            ifelse(ma==fa, 'either', 'lower')),
         correct = ifelse(ma!=fa, response==correct_response, response),
         correct = ifelse(correct=='upper', TRUE, correct),
         correct = ifelse(correct=='lower', FALSE, correct)) %>%
  group_by(level, ma, fa, correct) %>%
  nest() %>%
  mutate(mean = map(data, ~mean(.x$rt)),
         sd = map(data, ~sd(.x$rt)),
         skew = map(data, ~skewness(.x$rt))) %>%
  rowwise() %>%
  mutate(stats=paste('mean=',sprintf(fmt='%.1f',mean),
                     '\nsd=',sprintf(fmt='%.1f',sd),
                     '\nskew=',sprintf(fmt='%.1f',skew$Skewness), sep='')) %>%
  ungroup() %>%
  unnest(data) %>%
  ungroup()

sampled_fit_dist.plot = sampled_fit_dist %>%
  mutate(mafa=paste('mAdv=',ma,', ','fAdv=',fa,sep='')) %>%
  mutate(mafa=factor(mafa, levels=c('mAdv=1, fAdv=1','mAdv=2, fAdv=2',
                                    'mAdv=2, fAdv=1','mAdv=1, fAdv=2',
                                    'mAdv=3, fAdv=1','mAdv=1, fAdv=3'))) %>%
  mutate(response = as.character(response)) %>%
  mutate(correct_response = ifelse(ma>fa, 'upper',
                            ifelse(ma==fa, 'either', 'lower')),
         correct = ifelse(ma!=fa, response==correct_response, response),
         correct = ifelse(correct=='upper', TRUE, correct),
         correct = ifelse(correct=='lower', FALSE, correct)) %>%
  group_by(level, ma, fa, correct) %>%
  nest() %>%
  mutate(mean = map(data, ~mean(.x$rt)),
         sd = map(data, ~sd(.x$rt)),
         skew = map(data, ~skewness(.x$rt))) %>%
  rowwise() %>%
  mutate(stats=paste('mean=',sprintf(fmt='%.1f',mean),
                     '\nsd=',sprintf(fmt='%.1f',sd),
                     '\nskew=',sprintf(fmt='%.1f',skew$Skewness), sep='')) %>%
  ungroup() %>%
  unnest(data) %>%
  ungroup()
```

```{r, fig.width=4.2, fig.height=4.8}
bin_width = 0.06
fit_scale_factor = bin_width*nrow(data.fit.plot)/nrow(sampled_fit_dist.plot)

# trial = 'base'
# p.ddm.rt.base = ggplot() +
trial = 'subgoal'
p.ddm.rt.subgoal = ggplot() +
    geom_histogram(data=data.fit.plot %>% filter(level==trial, correct==TRUE),
                   mapping=aes(x=rt),
                   binwidth=bin_width,
                   color='white', size=.2,
                   fill=redblue[5]) +
    geom_histogram(data=data.fit.plot %>% filter(level==trial, correct==FALSE),
                     mapping=aes(x=rt, y =-..count..), # plot on reverse side of xaxis
                     binwidth=bin_width,
                     color='white', size=.2,
                     fill=redblue[5]) +
    geom_text(data=data.fit.plot %>% filter(level==trial, correct==TRUE) %>%
                group_by(mafa,stats) %>% nest() %>% ungroup(),
              mapping=aes(label=stats),
              x=.5, y=10, hjust=0, vjust=0,
              color=redblue[5], size=2.5, lineheight=.7) +
    geom_text(data=data.fit.plot %>% filter(level==trial, correct==FALSE) %>%
                group_by(mafa,stats) %>% nest() %>% ungroup(),
              mapping=aes(label=stats),
              x=.5, y=-25, hjust=0, vjust=0,
              color=redblue[5], size=2.5, lineheight=.7) +
    geom_density(data=sampled_fit_dist.plot %>% filter(level==trial, correct==TRUE),
                 mapping=aes(x=rt, y=..count..*fit_scale_factor),
                 color=redblue[3],
                 alpha=.0) +
    geom_density(data=sampled_fit_dist.plot %>% filter(level==trial, correct==FALSE),
                   mapping=aes(x=rt, y=-..count..*fit_scale_factor),
                   color=redblue[3],
                   alpha=.0) +
    geom_text(data=sampled_fit_dist.plot %>% filter(level==trial, correct==TRUE) %>%
                     group_by(mafa,stats) %>% nest() %>% ungroup(),
              mapping=aes(label=stats),
              x=1., y=10, hjust=0, vjust=0,
              color=redblue[3], size=2.5, lineheight=.7) +
    geom_text(data=sampled_fit_dist.plot %>% filter(level==trial, correct==FALSE) %>%
                     group_by(mafa,stats) %>% nest() %>% ungroup(),
              mapping=aes(label=stats),
              x=1., y=-25, hjust=0, vjust=0,
              color=redblue[3], size=2.5, lineheight=.7) +
    scale_x_log10(name='z(First Move Response Time)') +
    scale_y_continuous(name='N(trials)', labels=abs) +
    facet_wrap(.~mafa, ncol=2) +
    coord_cartesian(ylim=c(-30,60), # -50,100 for full data, -30, 60 for top/bottom
                    xlim=c(0.4,30))
```

# summarize weight estimates
```{r}
xfull = paired_fits %>% filter(model=='basemodel_dpsg_gd')
xtop = paired_fits
xbot = paired_fits

xfull = xfull %>% select(md, fd, gd, dpsg) %>% mutate(group='full')
xtop = xtop %>% select(md, fd, gd, dpsg) %>% mutate(group='tophalf')
xbot = xbot %>% select(md, fd, gd, dpsg) %>% mutate(group='bottomhalf')

xbase = xtop %>%
  rbind(xbot) %>%
  select(md, fd, group) %>%
  pivot_longer(c('md','fd'), names_to='param', values_to='value') %>%
  mutate(param=ifelse(param=='md', 'Start', 'Goal')) %>%
  mutate(group=ifelse(group=='tophalf', 'More Accurate', 'Less Accurate')) %>%
  mutate(param=factor(param, levels=c('Start','Goal')),
         group=factor(group, levels=c('More Accurate','Less Accurate'))) %>%
  mutate(level='base')
  
xsubgoal = xtop %>%
  rbind(xbot) %>%
  mutate(mdsg=md*dpsg, fdsg=fd*dpsg) %>%
  select(mdsg, fdsg, gd, group) %>%
  pivot_longer(c('mdsg','fdsg','gd'), names_to='param', values_to='value') %>%
  mutate(param=ifelse(param=='mdsg', 'Start', 
                      ifelse(param=='fdsg', 'Subgoal', 'Final\nGoal'))) %>%
  mutate(group=ifelse(group=='tophalf', 'More Accurate', 'Less Accurate')) %>%
  mutate(param=factor(param, levels=c('Start','Subgoal','Final\nGoal')),
         group=factor(group, levels=c('More Accurate','Less Accurate'))) %>%
  mutate(level='subgoal')

pleft = ggplot() +
  stat_summary(data=xbase, 
               mapping=aes(x=param, y=value, group=group, color=group, fill=group),
               fun='mean',
               geom='bar',
               width=0.8,
               position=position_dodge(width=.9)) +
  stat_summary(data=xbase,
               mapping=aes(x=param, y=value, group=group),
               fun.data='mean_cl_boot',
               geom='errorbar',
               position=position_dodge(width=.9)) +
  scale_color_manual(values=c('black', 'black')) +
  scale_fill_manual(values=c('white', 'gray')) +
  labs(y='Weight', x='', fill='Participant Group', color='Participant Group') +
  scale_y_continuous(limits=c(0,2.2), breaks=c(0,1,2)) +
  geom_text(mapping=aes(x=1, y=2.2, label='Base Task'),
            hjust=0, size=4, color='black') +
  theme(legend.position='none')

pright = ggplot() +
  stat_summary(data=xsubgoal, 
               mapping=aes(x=param, y=value, group=group, color=group, fill=group),
               fun='mean',
               geom='bar',
               width=0.8,
               position=position_dodge(width=.9)) +
  stat_summary(data=xsubgoal,
               mapping=aes(x=param, y=value, group=group),
               fun.data='mean_cl_boot',
               geom='errorbar',
               position=position_dodge(width=.9)) +
  scale_color_manual(values=c('black', 'black')) +
  scale_fill_manual(values=c('white', 'gray')) +
  labs(y='Weight', x='', fill='Participant Group', color='Participant Group') +
  scale_y_continuous(limits=c(0,2.2), breaks=c(0,1,2)) +
  geom_text(mapping=aes(x=1.25, y=2.2, label='Complex Task'),
            hjust=0, size=4, color='black') +
  theme(axis.line.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        legend.position=c(0, 0.8))

pweight = pleft + pright +
  plot_layout(ncol=2, widths=c(0.75,1))
```

# combined figure for IA trials
```{r, fig.width=3.4, fig.height=2.6}
( p.ia.main | p.model.comp.full) / (p.model.params | pweight) +
  plot_layout(nrow=2, heights=c(1,1)) &
  plot_annotation(tag_levels=list(c('A','B','C','','','D'))) &
  theme(axis.title = element_text(size=10),
        legend.key.size=unit(0.6, 'line'),
        legend.title=element_text(size=8),
        legend.text=element_text(size=8),
        legend.background = element_blank())

# ggsave(paste(plotout,'exp2_ia.pdf', sep=''), bg='transparent')
```

```{r, fig.width=7.5, fig.height=4.2}
p.ddm.rt.base + p.ddm.rt.subgoal +
  plot_annotation(tag_levels=c('A')) +
  plot_layout(ncol=2, widths=c(1,1)) &
  theme(axis.title = element_text(size=10))

# ggsave(paste(plotout,'exp2_ia_ddmrt.pdf', sep=''), bg='transparent')
```

```{r, fig.width=11, fig.height=3.6}
p.model.params.tophalf = p.model.params
p.ddm.rt.base.tophalf = p.ddm.rt.base
p.ddm.rt.subgoal.tophalf = p.ddm.rt.subgoal

p.model.params.tophalf | p.ddm.rt.base.tophalf | p.ddm.rt.subgoal.tophalf +
  plot_annotation(tag_levels=c('A', '1')) &
  theme(axis.title = element_text(size=10),
        legend.key.size=unit(0.6, 'line'),
        legend.title=element_text(size=8),
        legend.text=element_text(size=8),
        legend.background = element_blank())

# ggsave(paste(plotout,'exp2_ia_si_top.pdf', sep=''), bg='transparent')

p.model.params.bothalf = p.model.params
p.ddm.rt.base.bothalf = p.ddm.rt.base
p.ddm.rt.subgoal.bothalf = p.ddm.rt.subgoal

p.model.params.bothalf | p.ddm.rt.base.bothalf | p.ddm.rt.subgoal.bothalf +
  plot_annotation(tag_levels=c('A', '1')) &
  theme(axis.title = element_text(size=10),
        legend.key.size=unit(0.6, 'line'),
        legend.title=element_text(size=8),
        legend.text=element_text(size=8),
        legend.background = element_blank())

# ggsave(paste(plotout,'exp2_ia_si_bottom.pdf', sep=''), bg='transparent')
```

## survey
``` {r}
data.survey = data.survey %>%
  mutate_at(c('q1', 'q3', 'q5', 'q6', 'q7', 'q8', 'q9', 'q10'), 
            ~factor(., levels=c('never','rarely','sometimes','often','always')))
```

## likert questions
```{r, fig.height=2.5, fig.width=1.8}

plotdata = data.survey %>%
  select('subject', 'q1', 'q3', 'q5', 'q6', 'q7', 'q8', 'q9') %>%
  mutate_at(vars(starts_with('q')), as.numeric) %>%
  pivot_longer(c('q1', 'q3', 'q5', 'q6', 'q7', 'q8', 'q9'),
               names_to = 'question',
               values_to = 'likert') %>%
  separate(question, into=c('bla','question'), sep='q') %>%
  mutate(question = as.numeric(question)) %>%
  select(-bla) %>%
  mutate(question=factor(question, levels=10:1))

plotdata %>%
  ggplot(data=.,
         mapping=aes(x=likert, y=question, color=question, fill=question)) +
  ggridges::geom_density_ridges(stat='binline', alpha=0.2) +
  # geom_vline(data=surveymeans, mapping=aes(xintercept=likert, color=question),
  #            size=0.3) +
  scale_x_continuous(limits=c(0.8,5.2),
                     breaks=1:5) +
  labs(x='Likert Scale', y='Question No.') +
  theme(axis.title = element_text(size=10),
        legend.position = 'none')

# ggsave(paste(plotout,'survey_hist.pdf', sep=''))

plotdata %>%
  ggplot(data=.,
         mapping=aes(x=question, y=likert, color=question, fill=question)) +
  stat_summary(fun.data = "mean_cl_boot",
             geom = "pointrange",
             shape = 21,
             size = .6) +
  scale_y_continuous(limits=c(0.8,5.2),
                     breaks=1:5) +
  coord_flip() +
  labs(x='Question', y='Likert Scale') +
  theme(axis.title = element_text(size=10),
        legend.position = 'none')

# ggsave(paste(plotout,'survey_mean.pdf', sep=''))
```

## wordcloud util function
```{r}
rquery.wordcloud <- function(x, type=c("text", "url", "file"), 
                          lang="english", excludeWords=NULL, 
                          textStemming=FALSE,  colorPalette="Dark2",
                          min.freq=3, max.words=200) { 
  library("tm")
  library("SnowballC")
  library("wordcloud")
  library("RColorBrewer") 
  
  if(type[1]=="file") text <- readLines(x)
  else if(type[1]=="url") text <- html_to_text(x)
  else if(type[1]=="text") text <- x
  
  # Load the text as a corpus
  docs <- Corpus(VectorSource(text))
  # Convert the text to lower case
  docs <- tm_map(docs, content_transformer(tolower))
  # Remove numbers
  docs <- tm_map(docs, removeNumbers)
  # Remove stopwords for the language 
  docs <- tm_map(docs, removeWords, stopwords(lang))
  # Remove punctuations
  docs <- tm_map(docs, removePunctuation)
  # Eliminate extra white spaces
  docs <- tm_map(docs, stripWhitespace)
  # Remove your own stopwords
  if(!is.null(excludeWords)) 
    docs <- tm_map(docs, removeWords, excludeWords) 
  # Text stemming
  if(textStemming) docs <- tm_map(docs, stemDocument)
  # Create term-document matrix
  tdm <- TermDocumentMatrix(docs)
  m <- as.matrix(tdm)
  v <- sort(rowSums(m),decreasing=TRUE)
  d <- data.frame(word = names(v),freq=v)
  # check the color palette name 
  if(!colorPalette %in% rownames(brewer.pal.info)) colors = colorPalette
  else colors = brewer.pal(8, colorPalette) 
  # Plot the word cloud
  set.seed(1234)
  wordcloud(d$word,d$freq, min.freq=min.freq, max.words=max.words,
            random.order=FALSE, rot.per=0, 
            use.r.layout=TRUE, colors=colors)
  
  invisible(list(tdm=tdm, freqTable = d))
}
```

## q4 wordcloud
``` {r}
q4.answers = paste(data.survey %>% filter(!is.na(totalBonus)) %>% pull(q4), 
                   collapse = ' ')

rquery.wordcloud(q4.answers, type ="text", lang = "english",
                 colorPalette = "RdBu")
```

## q11 wordcloud
``` {r}
q11.answers = paste(data.survey %>% filter(!is.na(totalBonus)) %>% pull(q11), 
                   collapse = ' ')

rquery.wordcloud(q11.answers, type ="text", lang = "english",
                 colorPalette = "RdBu")
```

# session info
```{r}
sessionInfo()
```